---
title: "Final-Project"
author: "Annie Georgeson"
format: pdf
editor: visual
---

Loading in the libraries:

```{r}
library(tidyverse)
library(ggthemes)
library(dplyr)

```

Loading in the data: 

```{r}
food_insecurity = read_csv(here::here("2018_food_insecurity_data.csv"))
census = read_csv(here::here("clean_ar_census_2018.csv"))
income = read_csv(here::here("ar_income_data.csv"))

```

Filtering for just Arkansas's food insecurity data:

```{r}
arkansas_food = food_insecurity |>
  filter(State %in% c("AR"))

```

Joining the food insecurity data to census data:

```{r}
food_census_join = left_join(census, arkansas_food, by=c("Geographic Area Name" = "County, State"))

```

Joining the income data as well:
```{r}
food_census_join = left_join(food_census_join, income, by=c("Geographic Area Name" = "Geographic Area Name"))

```

Creating a new column that represents the percent of people per county that are food insecure:

```{r}
food_census_join = food_census_join |>
  mutate(percent_food_insecure = (`# of Food Insecure Persons in 2018`/ `Estimate!!SEX AND AGE!!Total population` * 100))

```


Identifying the counties with the highest & lowest percentage of food insecure persons:

```{r}
max_value = max(food_census_join$percent_food_insecure, na.rm = TRUE)
min_value = min(food_census_join$percent_food_insecure, na.rm = TRUE)

max_value #Phillips County
min_value #Benton County

```

Creating a new column that represents what percentage of residents are white in each respective county: 

```{r}
food_census_join = food_census_join |>
  mutate(percent_white_population = (`Estimate!!RACE!!Total population!!One race!!White`/ `Estimate!!SEX AND AGE!!Total population` * 100))

```

Identifying which counties have the highest & lowest percentage of white residents: 

```{r}
max_value_white = max(food_census_join$percent_white_population, na.rm = TRUE)
min_value_white = min(food_census_join$percent_white_population, na.rm = TRUE)

max_value_white #Cleburne County
min_value_white #Phillips County
```

Plot to represent the relationship between the percentage of white residents and the percent of the county considered food insecure: 

``` {r}
ggplot(food_census_join, aes(x=percent_white_population, y=percent_food_insecure)) +
  xlab("Percentage of White Residents per County") +
  ylab("Percentage of Food Insecurity per County") + 
  geom_point(size=.5)

```

Calculating what percentage of the total income of Arkansas was considered below the poverty line (<$20,780):

```{r}
food_census_join = food_census_join |>
  mutate(percent_poverty_line = (`Estimate!!Households!!Total!!Less than $10,000` + `Estimate!!Households!!Total!!$10,000 to $14,999` + `Estimate!!Households!!Total!!$15,000 to $24,999`))

```

Identifying what counties have the highest and lowest poverty rates: 

```{r}
max_value_poverty = max(food_census_join$percent_poverty_line, na.rm = TRUE)
min_value_poverty = min(food_census_join$percent_poverty_line, na.rm = TRUE)

max_value_poverty # Lee County
min_value_poverty # Saline County

```

Plot to represent the relationship between the percentage of households that fall below the poverty line and the percentage of food insecure populations: 

```{r}
ggplot(food_census_join, aes(x=percent_poverty_line, y=percent_food_insecure)) +
  geom_point(size=0.5) +
  xlab("Percentage of the County Below the Poverty Line") +
  ylab("Percentage of Food Insecure Populations")

```

Running a regression model with multiple variables:

```{r}
multi_variable_model = lm(percent_food_insecure~percent_white_population+percent_poverty_line, food_census_join)

summary(multi_variable_model)

# treating percent_food_insecure as the dependent variable and percent_white_population & percent_poverty_line as the independent variables

```
